MaxEnt: What is it good for?

### 2015-07-28 ms.tex
New sections
Working on the intro

\section{Introduction}
The principle of maximum entropy (MaxEnt) is an idea in probability theory
that is usually associated with \citet{jaynes}, although it has appeared in
various guises in the work of others \citep[e.g.][]{gibbs, boltzmann, shannon}.

The principle may be used
to {\it assign} or {\it update} a
subjective\footnote{in the sense that the probability distribution is a model
of the state of uncertainty of an idealised rational agent, rather than
a {\it frequency distribution} such as a histogram of people's heights.}
probability distribution when you obtain a certain kind of information
called {\it testable information}. In this article I will focus on the
updating application, as the assigning application can be viewed as a special
case where the prior distribution happened to be
uniform\footnote{A uniform probability distribution, either over a discrete
hypothesis space, or a continuous one with respect to a specific choice of
parameterisation, may be justified by a symmetry argument, or simply asserted
as in the ``subjective Bayesian'' approach.}.

The principle is very intuitively compelling
and there are some axiomatic arguments in its favour
\citep[e.g.][]{shore_johnson, 2006AIPC..872...31C, 2010arXiv1008.4831K}
and counterexamples to alternatives \citep[e.g.][]{presse}.
However, MaxEnt has also been widely
misapplied (even by some of its staunchest defenders), misunderstood,
and poorly defended (at times, even by Jaynes himself).
Consequently, most modern Bayesians pay little attention to the principle,
although there is a tendency among physicists to invoke it as a justification
for a prior distribution (be it uniform, exponential, gaussian, or whatever).
The purpose of this article is to explore the principle from the point of
view of a Bayesian practitioner, discuss some of the properties that
make MaxEnt compelling, and propose some ideas about how it could be used in
practice.

@@

\section{MaxEnt with constraints on marginals: ``hierarchical Bayes''}
Suppose we have a discrete prior distribution $q(\xx)$. Consider some
function of the unknown quantity, $f(\xx)$.
If the value of $f(\xx)$ is known, then our distribution for $\xx$ can
be updated from $q$ to $p$ as in Section~\ref{sec:bayes}.
If we receive testable information
about the expected value of $F$, then our distribution for $\xx$ can
be updated as in Section~\ref{sec:expectations}.

@@

>>>>

\section{Introduction}
It is now well established that degrees of certainty can and should be modelled
with probability theory. The resulting framework is usually called Bayesian
inference, whose key feature is the updating of probabilities to incorporate
new information when it is learned.

The principle of maximum entropy (MaxEnt) is an idea that was developed...

It is usually associated with \citet{jaynes}, although it has appeared in
various guises in the work of others \citep[e.g.][]{gibbs, boltzmann, shannon}.

MaxEnt can be used either to {\it assign} or to {\it update} a
subjective\footnote{in the sense that the probability distribution is a model
of the state of uncertainty of an idealised rational agent, rather than
a {\it frequency distribution} such as a histogram of people's heights.}
probability distribution when you obtain a certain kind of information
called {\it testable information}. In this article I will focus on the
updating application, as the assigning application can be viewed as a special
case where the prior distribution happened to be
uniform\footnote{A uniform probability distribution, either over a discrete
hypothesis space, or a continuous one with respect to a specific choice of
parameterisation, may be justified by a symmetry argument, or simply asserted
as in the ``subjective Bayesian'' approach.}.

There are some axiomatic arguments supporting MaxEnt
\citep[e.g.][]{shore_johnson, 2006AIPC..872...31C, 2010arXiv1008.4831K},
as well as counterexamples to possible alternatives \citep[e.g.][]{presse}.
However, there is no consensus among Bayesians of the status of MaxEnt: is
it a valid and potentially useful fundamental principle, some kind of an
approximation or computational trick, or a dead end that is solely of
historical interest?

has also been widely
misapplied (even by some of its staunchest defenders), misunderstood,
and poorly defended (at times, even by Jaynes himself).
Consequently, most modern Bayesians pay little attention to the principle,
although there is a tendency among physicists to invoke it as a justification
for a prior distribution (be it uniform, exponential, gaussian, or whatever).
The purpose of this article is to explore the principle from the point of
view of a Bayesian practitioner, discuss some of the properties that
make MaxEnt compelling, and propose some ideas about how it could be used in
practice.

@@

\section{MaxEnt with constraints on marginals}
\subsection{Heirarchical Bayes}
Suppose we have a discrete prior distribution $q(\xx)$. Consider some
function of the unknown quantity, $f(\xx)$.
If the value of $f(\xx)$ is known, then our distribution for $\xx$ can
be updated from $q$ to $p$ as in Section~\ref{sec:bayes}.
If we receive testable information
about the expected value of $F$, then our distribution for $\xx$ can
be updated as in Section~\ref{sec:expectations}.

@@

\subsection{Jeffrey Conditionalisation}
